Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2023-04-28 paper-correlations.tex
Update pore size function

Before computation of chord length function an input array must be pre-processed
to label an interface for the phase of interest. For this purpose we use
distance transform function defined as follows:
\begin{equation}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right. \label{eq:distance-transform}
\end{equation}
Distance transform has many implementations \textcolor{red}{cite something},
many of them can be parralelized on GPU. Array pre-processing is performed in
\code{cl\_preprocess} function and per-slice chord lengths accumulation is
performed in \code{cl\_slice} function. After all cord lengths are accumulated
they can be binned in a histogram. In the following algorithm we suppose that
phases in the input are labeled with non-negative integers.
\begin{algorithmic}[1]
  \Procedure{chord\_length}{$A, phase$}
    \State $A' \gets cl\_preprocess(A, phase)$
    \State $chordlist \gets [\ ]$
    \ForAll{$slice \in slices(A')$}
      \State Append $cl\_slice(A', phase)$ to $chordlist$
    \EndFor
    \State \textbf{return} $chordlist$ or $histogram(chordlist)$  
  \EndProcedure
  \\
  \Procedure{cl\_preprocess}{$A, phase$}
    \State $A' \gets I^{(phase)}(A)$
    \State $edge\_label \gets maximum(A) + 1$
    \Comment Choose an unused label to mark elements of the interface with
    \State $A_D = \mathcal{D}(A')$
    \Comment Perform distance transform for all elements of $A'$
    \State $E \gets edge\_label \cdot (A_D = 1)$
    \Comment We consider elements of $A_D$ equal to $1$ to be part of the
    interface and mark them so
    \State $R \gets min(A + E, edge\_label)$
    \Comment Add interface labels to the input
    \State \textbf{return} $R$
  \EndProcedure
  \\
  \Procedure{cl\_slice}{$A, phase$}
    \State $chordlist \gets [\ ]$
    \State $len \gets 0$
    \State $maybechord \gets false$
    \State $edge\_label \gets maximum(A)$
    \ForAll{$x \in A$}
      \If{$x = edge\_label$}
        \If{$len > 1\ and\ maybechord$}
          \State $chordlist \gets (len-1):chordlist$
          \Comment A chord with length $len-1$ voxels is found
        \EndIf
        \State $len \gets 0$
        \State $maybechord \gets true$
      \ElsIf{$x \ne phase$}
        \State $maybechord \gets false$
      \EndIf
      \State $len \gets len+1$
    \EndFor
    \State \textbf{return} $chordlist$
    \EndProcedure
\end{algorithmic}

Another function which does not use this approach but is in the
\code{Directional} module for historical reasons is the pore size function
$P(r)$. Remember that pore size function is a probability distribution
function. In practice it is better represented as a histogram of pore sizes
found in the input. In our approach we calculate this histogram using distance
transform. For each point $\bm{x}$ distance transform $\mathcal{D}(\bm{x})$ is
defined as
\begin{equation*}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right.
\end{equation*}
Distance transform has many implementations \textcolor{red}{cite something}. Our
algorithm for calculating $P(r)$ is then:
\begin{algorithmic}[1]
  \Procedure{$P$}{$A, nbins$}
    \State $D \gets \mathcal{D}(A)$ \Comment{Apply $\mathcal{D}$ to each element of $A$}
    \State $D' = \{ x \in D: x \ne 0\}$
    \Comment Remove zeros
    \State Bin values in $D'$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

>>>>

Before computation of chord length function an input array must be pre-processed
to label an interface for the phase of interest. For this purpose we use
distance transform function defined as follows:
\begin{equation}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right. \label{eq:distance-transform}
\end{equation}
Distance transform has many implementations \textcolor{red}{cite something},
many of them can be parralelized on GPU. Array pre-processing is performed in
\code{cl\_preprocess} function and per-slice chord lengths accumulation is
performed in \code{cl\_slice} function. After all cord lengths are accumulated
they can be binned in a histogram. In the following algorithm we suppose that
phases in the input are labeled with non-negative integers.
\begin{algorithmic}[1]
  \Procedure{chord\_length}{$A, phase$}
    \State $A' \gets cl\_preprocess(A, phase)$
    \State $chordlist \gets [\ ]$
    \ForAll{$slice \in slices(A')$}
      \State Append $cl\_slice(A', phase)$ to $chordlist$
    \EndFor
    \State \textbf{return} $chordlist$ or $histogram(chordlist)$  
  \EndProcedure
  \\
  \Procedure{cl\_preprocess}{$A, phase$}
    \State $A' \gets I^{(phase)}(A)$
    \State $edge\_label \gets maximum(A) + 1$
    \Comment Choose an unused label to mark elements of the interface with
    \State $A_D = \mathcal{D}(A')$
    \Comment Perform distance transform for all elements of $A'$
    \State $E \gets edge\_label \cdot (A_D = 1)$
    \Comment We consider elements of $A_D$ equal to $1$ to be part of the
    interface and mark them so
    \State $R \gets min(A + E, edge\_label)$
    \Comment Add interface labels to the input
    \State \textbf{return} $R$
  \EndProcedure
  \\
  \Procedure{cl\_slice}{$A, phase$}
    \State $chordlist \gets [\ ]$
    \State $len \gets 0$
    \State $maybechord \gets false$
    \State $edge\_label \gets maximum(A)$
    \ForAll{$x \in A$}
      \If{$x = edge\_label$}
        \If{$len > 1\ and\ maybechord$}
          \State $chordlist \gets (len-1):chordlist$
          \Comment A chord with length $len-1$ voxels is found
        \EndIf
        \State $len \gets 0$
        \State $maybechord \gets true$
      \ElsIf{$x \ne phase$}
        \State $maybechord \gets false$
      \EndIf
      \State $len \gets len+1$
    \EndFor
    \State \textbf{return} $chordlist$
    \EndProcedure
\end{algorithmic}
This implementation also works in linear time.

Another function built on top of distance transform is the pore size function
$P(r)$. Remember that pore size function is a probability distribution
function. In practice it is better represented as a histogram of pore sizes
found in the input. Our algorithm for calculating $P(r)$ is as follows:
\begin{algorithmic}[1]
  \Procedure{$P$}{$A$}
    \State $D \gets \mathcal{D}(A)$
    \Comment Apply $\mathcal{D}$ (\cref{eq:distance-transform}) to each element of $A$
    \State $D' = \{ x \in D: x \ne 0\}$
    \Comment Remove zeros
    \State Bin values in $D'$ to histogram $H$
    \State \textbf{return} $H$
    \Comment Alternatively just return $D'$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.
